### 2013-05-23 paper/main.tex
Initial commit

\documentclass[a4paper,10pt]{article}
\usepackage[british]{babel}
\usepackage[T1]{fontenc}
\usepackage{url}
\usepackage[a4paper]{geometry}

\usepackage{listings}
\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{float}
\usepackage{xspace}
\usepackage{enumerate}

% Math packages
\usepackage[sc]{mathpazo}
\usepackage{amsthm}
\usepackage{algorithmic}
\usepackage{amsmath}

\lstset{basicstyle=\footnotesize}

% Document properties
\title{Modelling and Analysis of an Elevator System with NuSMV2}
\author{
	Wouter Geraedts \\ \small{\texttt{w.geraedts@student.ru.nl}} \and
	Ko Stoffelen     \\ \small{\texttt{kostoffelen@student.ru.nl}}
}
\date{tbd}
\linespread{1.05}

% Terminology
\newcommand{\NuSMV}{NuSMV2\xspace}

% Math
\newcommand{\LTLG}{\mathbf{G~}}
\newcommand{\LTLF}{\mathbf{F~}}
\newcommand{\LTLX}{\mathbf{X~}}
\newcommand{\LTLU}{\mathbf{~U~}}
\newcommand{\disjall}[1]{\mathop{\bigvee}\limits_{i=0}^{#1}}
\newcommand{\conjall}[2]{\mathop{\bigwedge}\limits_{#1}^{#2}}
\newcommand{\dooropen}[1]{\mathsf{door\_open}(#1)}
\newcommand{\cabinat}[1]{\mathsf{cabin\_at}(#1)}
\newcommand{\requested}[1]{\mathsf{requested}(#1)}
\newcommand{\served}[1]{\mathsf{served}(#1)}
\newcommand{\imply}{\rightarrow}

\begin{document}
	\maketitle

	\section{Introduction}

	% Baier & katoen 5.5
	% LTL

	\section{Problem 1}

	\quote{Work out the solution to Exercise 5.5 of Baier \& Katoen (page 302/303)}

	\begin{enumerate}[(a)]
		\item
			The doors are ``safe'', i.e., a floor door is never open if the elevator is not present at the given floor.

			\[
				\LTLG \neg \left( \conjall{i=0}{N-1} \dooropen{i} \wedge \neg \cabinat{i} \right)
			\]

		\item
			A requested floor will be served sometime.

			\begin{align*}
				\served{x} := \cabinat{x} \wedge \dooropen{x} \\
				\LTLG \conjall{i=0}{N-1} \requested{i} \imply \LTLF \served{i}
			\end{align*}

		\item
			Again and again the elevator returns to floor 0.

			\[
				\LTLG \LTLF \cabinat{0}
			\]

		\item
			When the top floor is requested, the elevator serves it immediately and does not stop on the way there.

			\[
				\LTLG \requested{N-1} \imply
					\left(
						\conjall{i=0}{N-2}
							\neg \dooropen{i} \wedge ( \cabinat{i} \imply \LTLX \cabinat{i+1} )
					\right)
					\LTLU
					\served{N-1}
			\]
	\end{enumerate}

	\bibliographystyle{plain}
	\bibliography{main}

\end{document}